---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
### Setup
```{r}
rm(list=ls())
library(readr)
library(ggplot2)
library(lattice)
library(knitr)
options(scipen = 1, digits = 2)
```

### Loading and preprocessing the data
```{r}
raw_data=read.csv(unzip("activity.zip"))
raw_data$interval<-factor(raw_data$interval, levels=unique(raw_data$interval))
df<-raw_data[complete.cases(raw_data),]

```


### Mean total number of steps taken per day
```{r}
total_per_day<-aggregate(x=df$steps, by = list(df$date), sum)
hist(total_per_day$x, main="Mean Total Steps per Day",xlab="Steps per day",col="grey")
```
###Medium and Mean
```{r}
rmean <- mean(total_per_day$x)
rmed<- median(total_per_day$x)
```

The MEAN is **`r rmean`**, and MEDIUM is **`r rmed`**

### The average daily activity pattern?
``` {r}
steps_per_interval<-aggregate(x=df$steps, by = list(df$interval), mean)
plot(steps_per_interval$Group.1, steps_per_interval$x, type="l", xlab= "Interval", ylab="Steps", main="Average Number of Steps per Day by Interval")
lines(steps_per_interval$Group.1, steps_per_interval$x, type="l")
```

### The 5-minute interval containing the maximum number of steps?
```{r}
total_per_interval<-aggregate(x=df$steps, by = list(df$interval), sum)
max_steps<-total_per_interval[which.max((total_per_interval$x)),1] 

```

The Interval with Maximum steps is **`r max_steps`**

### Imputing missing values



```{r}
na_rows<-raw_data[is.na(raw_data$steps),]    

```

The total number of rows with NAs is **`r nrow(na_rows)`**

```{r}
#Update missing values with Category-Mean
#-calculate mean for each category, ignoring missing values
lookup_table <- tapply(raw_data$steps, raw_data$interval, mean, na.rm=TRUE) 
#-get the index of the missing values
na_position<-which(is.na(raw_data$steps))
#-newdata with NAs replaced by interval mean
new_data<-raw_data 
new_data$steps[na_position]<- lookup_table[raw_data$interval[na_position]] #using index on the lookup_table to acquire appropriate mean and update the missingh values!

#-Histogram of the total number of steps taken each day
new_total_per_day<-aggregate(x=new_data$steps, by = list(new_data$date), sum)
hist(new_total_per_day$x, main="Mean Total Steps per Day",xlab="Steps",col="darkgreen")

```
```{r}
#- Mean and Medium
new_mean<-mean(new_total_per_day$x)
new_median<-median(new_total_per_day$x)
```
The imputed mean is **`r new_mean`**
The imputed median is **`r new_median`**


### Activity patterns between weekdays and weekends?

```{r}
new_data$date<-as.Date(new_data$date)
new_data$day_type<-ifelse(weekdays(new_data$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
new_data <- aggregate(steps ~ interval + day_type, new_data, mean)
xyplot(new_data$steps ~ new_data$interval|new_data$day_type, main="Daily Average by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```